---
title: "`r params$title %||% 'Portfolio Analysis Report'`"
subtitle: "Quant Research & Economic Intelligence Platform"
author: "`r params$author %||% 'Quant Research Platform'`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    highlight: pygments
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    css: "../../assets/themes/report.css"
    self_contained: true
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: xelatex
params:
  cfg:    NULL
  title:  NULL
  author: NULL
  tickers: ["SPY", "AGG", "GLD", "QQQ", "TLT"]
  from:    "2020-01-01"
  to:      NULL
  rf:      0.0525
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  message     = FALSE,
  warning     = FALSE,
  fig.width   = 11,
  fig.height  = 6.5,
  dpi         = 180,
  fig.align   = "center",
  out.width   = "100%"
)

library(here)
source(here::here("R", "utilities", "config.R"))
source(here::here("R", "utilities", "helpers.R"))
source(here::here("R", "ingestion",  "yahoo_finance.R"))
source(here::here("R", "cleaning",   "clean_prices.R"))
source(here::here("R", "transformations", "portfolio_analytics.R"))
source(here::here("R", "visualization",   "theme.R"))
source(here::here("R", "visualization",   "charts.R"))

set_quant_theme()
cfg     <- if (!is.null(params$cfg)) params$cfg else get_config()
tickers <- params$tickers
rf      <- params$rf
to_date <- if (is.null(params$to)) Sys.Date() else as.Date(params$to)
```

---

> **Classification:** Internal Research
> **Horizon:** `r params$from` to `r format(to_date, "%Y-%m-%d")`
> **Benchmark:** S&P 500 (SPY)
> **Risk-Free Rate:** `r sprintf("%.2f%%", rf * 100)` (annualised)

---

## Executive Summary

```{r load-data}
prices_wide <- fetch_adjusted_prices(tickers, from = params$from, to = to_date, cfg = cfg)
prices_wide <- fill_price_gaps(prices_wide)

# Compute log returns
tickers_avail <- setdiff(names(prices_wide), "date")
returns_wide  <- prices_wide |>
  dplyr::mutate(across(-date, ~ c(NA_real_, log_returns(.x)))) |>
  dplyr::filter(!is.na(.data[[tickers_avail[1]]]))

# Equal-weight portfolio
n_assets <- length(tickers_avail)
weights  <- setNames(rep(1 / n_assets, n_assets), tickers_avail)
port     <- portfolio_statistics(returns_wide, weights, rf = rf)
```

This report provides a comprehensive quantitative analysis of a **`r n_assets`-asset portfolio** over the period **`r params$from`** to **`r format(to_date, "%Y-%m-%d")`** (`r port$stats["n_obs"]` trading days). The portfolio employs an equal-weight allocation, rebalanced continuously.

### Key Performance Indicators

```{r kpi-table}
s <- port$stats
kpi <- tibble::tibble(
  Metric         = c("Total Return", "Annualised Return", "Annualised Volatility",
                      "Sharpe Ratio", "Sortino Ratio", "Maximum Drawdown",
                      "Calmar Ratio", "VaR 95% (daily)", "Expected Shortfall 95%"),
  Value          = c(
    sprintf("%.2f%%", s["total_return"]     * 100),
    sprintf("%.2f%%", s["ann_return"]       * 100),
    sprintf("%.2f%%", s["ann_volatility"]   * 100),
    sprintf("%.3f",   s["sharpe_ratio"]),
    sprintf("%.3f",   s["sortino_ratio"]),
    sprintf("%.2f%%", s["max_drawdown"]     * 100),
    sprintf("%.3f",   s["calmar_ratio"]),
    sprintf("%.3f%%", s["var_95_daily"]     * 100),
    sprintf("%.3f%%", s["es_95_daily"]      * 100)
  ),
  Interpretation = c(
    "Cumulative compound return over full period",
    "CAGR using geometric compounding",
    "Annualised std dev × √252",
    "(Ann. Return − Rf) / Ann. Vol — target > 1.0",
    "(Ann. Return − Rf) / Downside Dev",
    "Peak-to-trough maximum loss",
    "Ann. Return / |Max Drawdown| — target > 1.0",
    "Daily loss not exceeded 95% of the time (normal)",
    "Expected loss given VaR is breached"
  )
)

knitr::kable(kpi, align = c("l","r","l"),
             caption = "Portfolio Performance Summary") |>
  kableExtra::kable_styling(bootstrap_options = c("striped","hover","condensed"),
                             full_width = FALSE)
```

---

## 1. Performance Analysis

### 1.1 NAV Trajectory

```{r plot-nav}
chart_performance(prices_wide,
  title    = "Portfolio Asset Performance",
  subtitle = "Adjusted closing prices, rebased to 100 at start of period")
```

*Chart reads: each line shows the cumulative return of a single asset if $100 were invested at the start of the period. Divergence between lines represents the cross-sectional dispersion of returns.*

### 1.2 Return Distributions

```{r return-dist, fig.height=8}
plots <- purrr::map(tickers_avail, function(t) {
  chart_return_distribution(returns_wide[[t]], ticker = t)
})
gridExtra::grid.arrange(grobs = plots, ncol = 2)
```

---

## 2. Risk Analysis

### 2.1 Drawdown (Underwater Chart)

```{r drawdown}
chart_drawdown(prices_wide)
```

### 2.2 Rolling 63-Day Volatility (Annualised)

```{r rolling-vol}
ret_long <- returns_wide |>
  tidyr::pivot_longer(-date, names_to = "ticker", values_to = "return")
chart_rolling_volatility(ret_long, window = 63)
```

---

## 3. Correlation & Covariance Structure

### 3.1 Correlation Matrix

```{r corr-matrix}
chart_correlation_matrix(returns_wide,
  title = "Pairwise Asset Correlation (Daily Log-Returns)")
```

### 3.2 Covariance Summary

```{r cov-table}
Sigma <- cov(returns_wide[, tickers_avail], use = "pairwise.complete.obs") * 252
round(Sigma * 100, 3) |>
  knitr::kable(caption = "Annualised Covariance Matrix (×100)",
               digits = 3) |>
  kableExtra::kable_styling(bootstrap_options = "striped")
```

---

## 4. CAPM & Factor Analysis

```{r capm}
capm_df <- capm_multi(returns_wide, benchmark_col = "SPY", rf_annual = rf)

knitr::kable(capm_df |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 4))),
  caption = "CAPM Results — OLS on Excess Returns vs S&P 500") |>
  kableExtra::kable_styling(bootstrap_options = c("striped","hover"))
```

**Interpretation:**

- **Alpha** > 0 indicates risk-adjusted outperformance of the benchmark (Jensen's α)
- **Beta** > 1 implies amplified market sensitivity (aggressive); < 1 implies defensiveness
- **R²** measures what fraction of return variation is explained by systematic (market) risk
- **Treynor Ratio** = (excess return) / beta — risk-adjusted return per unit of systematic risk

---

## 5. Portfolio Optimisation

### 5.1 Efficient Frontier

```{r efficient-frontier}
mu_vec  <- colMeans(returns_wide[, tickers_avail], na.rm = TRUE) * 252
Sigma_a <- cov(returns_wide[, tickers_avail], use = "complete.obs") * 252
frontier <- efficient_frontier(mu_vec, Sigma_a, n_points = 200, rf = rf)

asset_df <- tibble::tibble(
  ticker     = tickers_avail,
  ann_return = mu_vec,
  ann_vol    = sqrt(diag(Sigma_a))
)

chart_efficient_frontier(frontier, asset_df = asset_df, rf = rf)
```

---

## 6. Methodology Notes

### 6.1 Return Computation

All returns are computed as **log-returns**: $r_t = \ln(S_t / S_{t-1})$. Log-returns are preferred over simple returns because they are:

1. Time-additive: $r_{0,T} = \sum_{t=1}^T r_t$
2. Normally distributed (approximately) for short horizons
3. Symmetric around zero for gains/losses of equal magnitude

### 6.2 Risk Metrics

**Value at Risk (VaR):** Under the normal distribution assumption,
$$\text{VaR}_\alpha = -(\mu - z_\alpha \sigma)$$
where $z_\alpha = \Phi^{-1}(\alpha)$ is the $\alpha$-quantile of the standard normal.

**Expected Shortfall (ES):** The conditional expectation of loss beyond VaR,
$$\text{ES}_\alpha = -\left(\mu - \sigma \frac{\phi(z_\alpha)}{\alpha}\right)$$
where $\phi$ is the standard normal PDF. ES is a coherent risk measure; VaR is not.

### 6.3 Sharpe Ratio

$$\text{SR} = \frac{\mu_{\text{ann}} - r_f}{\sigma_{\text{ann}}}$$

Annualisation: $\mu_{\text{ann}} = (1 + \bar{r})^{252} - 1$, $\sigma_{\text{ann}} = \sigma_{\text{daily}} \times \sqrt{252}$.

---

## 7. Disclaimers

This report is generated by an automated quantitative research platform for **educational and research purposes only**. Past performance does not guarantee future results. No investment advice is implied or intended. All models carry assumptions that may not hold in practice.

---

*Generated by Quant Research & Economic Intelligence Platform v1.0.0*
*`r format(Sys.time(), "%Y-%m-%d %H:%M:%S UTC")`*
